---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

``` {r}
library(tidyverse)

deaths <- av %>%
  pivot_longer(cols = starts_with("Death"), names_to = "Time", values_to = "Death") %>%
  filter(Death != "") %>%
  mutate(Time = parse_number(Time))
```

Similarly, deal with the returns of characters.

``` {r}
returns <- av %>%
  pivot_longer(cols = starts_with("Return"), names_to = "Time", values_to = "Return") %>%
  filter(Return != "") %>%
  mutate(Time = parse_number(Time))

returns
```

Based on these datasets calculate the average number of deaths an Avenger suffers.
```{r}
total_deaths <- sum(deaths$Time)
total_returns <- sum(returns$Time)

num_avengers <- nrow(av)
average_deaths <- (total_deaths - total_returns) / num_avengers

average_deaths
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

Andrew- The MVP of the Earth-616 Marvel Universe Avengers has to be Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times.

mazin- Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.

Luke R- Given the Avengers’ 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so

There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.

### Include the code

```{r}
most_deaths_index <- which.max(deaths$Time)
character_most_deaths <- deaths$Name.Alias[most_deaths_index]
most_deaths <- deaths$Time[most_deaths_index]

most_returns_index <- which.max(returns$Time)
character_most_returns <- returns$Name.Alias[most_returns_index]
most_returns <- returns$Time[most_returns_index]

cat("Character with the most deaths:", character_most_deaths, "with", most_deaths, "deaths\n")
cat("Character with the most returns:", character_most_returns, "with", most_returns, "returns\n")
```


```{r}
avengers <- av %>%
  mutate(Total_Deaths = rowSums(select(., starts_with("Death")) == "YES"))


died_at_least_once <- sum(avengers$Total_Deaths > 0)


died_at_least_once

```


```{r}
death_year <- total_deaths / max(av$Years.since.joining)
death_period <- 1 / (death_year / 12)
death_period #Luke code
```


```{r}
num_died_once <- sum(av$Death1 == "YES" & av$Return1 == "YES") #number of people who died once
num_died_twice <- sum(av$Death2=="YES" & av$Return2 == "YES")
num_died_three <- sum(av$Death3=="YES" & av$Return3 == "YES")



cat("Percentage of a member from the avengers to return from 1st death" , num_died_once / sum(av$Death1 == "YES"))
print(" ")

cat("Percentage of a member from the avengers to return from 2nd death" , num_died_twice / sum(av$Death2 == "YES"))

print(" ")
cat("Percentage of a member from the avengers to return from 3rd death" , num_died_three / sum(av$Death3 == "YES"))

```

### Include your answer

Their statement was correct! Jocasta has the most deaths and returns with 5 each.

mazin- the statement was correct as well. after running it in all the deaths written with 173 listed avengers, it showed that the number of deaths more than once is 69.


Luke R- Our results are slightly off their statement. We have a death occurring every 6.2 months, but that is showing activity for 115 years, which doesn't match their statement of 53 years of activity.

Their statement was correct.

Upload your changes to the repository. Discuss and refine answers as a team.

